---
title: "EDUCATION IN TURKEY"
output: 
  html_document: 
    toc: yes
    df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<p>&nbsp;</p>


### **Import necessary libraries**

```{r message=FALSE, warning=FALSE}
library(readxl)
library(tidyverse)
library(dplyr)
library(funModeling)
library(reshape2)
library(magrittr)
library(sf)
library(ggplot2)
library(gganimate)
library(plotly)
library(qdap)
library(ggthemes)
```

<p>&nbsp;</p>



# **Population**

<p>&nbsp;</p>



### **Import the data**

```{r message=FALSE, warning=FALSE}
data <- read_excel("data/population.xls")
```

<p>&nbsp;</p>


### **Data manipulation**

```{r message=FALSE, warning=FALSE}
#Set column names as row 6 with distinct numbers to make ready to delete
#male and female columns
names(data) <- paste(data[6, ], seq(1,ncol(data)))


#Drop male and female columns
data <- select(data, -contains("Male"), -contains("Female"))


#Set names as row 4 in order to delete NA columns easily
names(data) <- data[4, ]


#Drop NA columns
data <- select(data, -starts_with("NA"))


#Get rows between 7 and 18 as last form of our dataset
data <- slice(data, 7:18)


#Rename columns with english expressions
names(data) <- c("year", "total", "illiterate", "literate_without_diploma",
                 "primary_school", "primary_education", "junior_and_vocational_high school",
                 "high_and_vocational_high_school", "universities_and_other_higher_educational_institutions",
                 "master", "doctorate", "unknown")


#Change dataframe type to numeric
data <- data.frame(lapply(data, function(x) as.numeric(as.character(x))))


```

<p>&nbsp;</p>


### **Let's have a look to our data**

```{r}
head(data)
```

<p>&nbsp;</p>


### **Structure of our data**

```{r}
str(data)
```

<p>&nbsp;</p>


### **Check if there are zeros, na's or infinitive**

```{r}
funModeling::status(data)
```

<p>&nbsp;</p>


### **Pie chart of education levels of population at each year**

```{r fig.height=10, fig.width=10, message=FALSE, warning=FALSE}
data_melted <- select(data, -total)
data_melted <- melt(data_melted, "year")

fig <- plot_ly(
  data_melted,
  labels = ~variable,
  values = ~value,
  type = 'pie',
  frame = ~year
  )

fig %<>% 
  layout(title = 'Population by Education Level',
  xaxis = list(showgrid = FALSE,
               zeroline = FALSE,
               showticklabels = FALSE),
  yaxis = list(showgrid = FALSE,
               zeroline = FALSE,
               showticklabels = FALSE),
  legend = list(x = 100,
                y = 0.5)
  )



fig
```

<p>&nbsp;</p>


### **Population by education level line plots**

```{r fig.height=6, fig.width=10, message=FALSE, warning=FALSE}
for (i in seq(2, 12)){
  data_melted <- melt(data, c(names(data)[-i]))
  scatter_plot <- ggplot(data_melted, aes(x = year, y = value)) + 
    geom_line(linetype = "longdash", lwd = 1) + 
    scale_x_continuous(breaks = data$year) +
    scale_y_continuous(labels = function(x) format(x, scientific = FALSE)) +
    theme_igray() +
    theme(strip.text = element_text(size = 10,
                                    face = "bold"),
          legend.text = element_blank(),
          legend.title = element_blank(),
          axis.text = element_text(size = 10),
          axis.title = element_text(size = 10,
                                    face = "bold"),
          axis.text.x = element_text(angle = 50,
                                     vjust = 0.5),
          axis.title.y = element_blank()) +
    facet_wrap(~variable,
               ncol = 1)
  
  
  print(scatter_plot)
}
```

<p>&nbsp;</p>
<p>&nbsp;</p>


# **Happiness Status of Different Education Levels**


<p>&nbsp;</p>


### **Import the data**

```{r message=FALSE, warning=FALSE}
data <- read_excel("data/hapiness.xls")
```

<p>&nbsp;</p>


### **Data manipulation**

```{r message=FALSE, warning=FALSE}
#Get words after "-" in order to obtain only english parts
for (column in names(data)){
  i = 1
  for (value in data[column]){
    data[column][i] <- gsub(".*-", "", value)
    i <- i + 1
  }
}


#Renaming variables
names(data) <- c("year", "level", "no_school_completed",  "primary_school",
                  "primary_education_or_junior_high_school", "high_school_or_equivalent",
                 "higher_education")


#Get dataset between 4-67 rows
data <- slice(data,  4:67)


#Shift year column by 1 row
data$"year" <- lag(data$"year", 1)


#Get only Happy rows and remove others
data <- filter(data, level == "Happy")


#We do not need level anymore
data$level <- NULL


#Change the type of values from character to numeric
data <- data.frame(lapply(data, function(x) as.numeric(as.character(x))))
```

<p>&nbsp;</p>


### **Have a look to our data**

```{r message=FALSE, warning=FALSE}
str(data)
```

<p>&nbsp;</p>


### **More data manipulation**

```{r message=FALSE, warning=FALSE}
#Transpose the dataframe and rename variables
data_transpose <- as.data.frame(t(as.matrix(data)))
names(data_transpose) <- data_transpose[1, ]
data_transpose <- (data_transpose[2:nrow(data_transpose), ])
data_transpose <- cbind(level = rownames(data_transpose), data_transpose)
names(data_transpose) <- paste("value", names(data_transpose), sep = "_")


#Change factor levels and date format in order to plot properly
data_transpose$value_level <- factor(data_transpose$value_level, levels = 
                                       c("no_school_completed",
                                         "primary_school",
                                         "primary_education_or_junior_high_school",
                                         "high_school_or_equivalent",
                                         "higher_education"))

head(data_transpose)
```

<p>&nbsp;</p>


### **Bar chart of happiness of different education groups**

```{r fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
data_melted <- melt(data, "year")
f1 <- list(
  size = 18
  )

f2 <- list(
  size = 14
  )

ay <- list(
  title = "HAPPINESS LEVEL",
  titlefont = f1,
  showticklabels = TRUE,
  tickfont = f2,
  exponentformat = "E"
  )

ax <- list(
  title = "",
  zeroline = FALSE,
  showline = FALSE,
  showticklabels = FALSE,
  showgrid = FALSE
  )

fig <- plot_ly(
  data_melted,
  x = ~variable,
  y = ~value,
  type = "bar",
  frame = ~year,
  color = ~variable,
  texttemplate = gsub("_", "<br>", data_melted$variable),
  textposition = 'outside'
  )

fig %<>% layout(xaxis = ax,
                yaxis = ay,
                showlegend = FALSE)
                uniformtext=list(minsize=10, mode = 'show')

fig
```

<p>&nbsp;</p>


### **Change of happiness over years line plot**

```{r fig.height=6, fig.width=13, message=FALSE, warning=FALSE}
data_melted <- melt(data, "year")
change_plot <- ggplot(data_melted, aes(x = year, y = value)) + 
  geom_line(aes(color = variable), lwd = 2) + 
  scale_x_continuous(breaks = data$year) +
  scale_color_manual(values = c("darkred", "steelblue", "green4", "purple3", "gray7")) +
  theme_bw()+  
  theme(legend.text = element_text(size = 10),
        legend.title = element_blank(),
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 15, face = "bold"),
        axis.text.x = element_text(angle = 50, vjust = 0.5),
        axis.title.y = element_blank())

change_plot
```

<p>&nbsp;</p>
<p>&nbsp;</p>


# **Income Status of Different Education Levels**

<p>&nbsp;</p>


### **Import the dataset**

```{r message=FALSE, warning=FALSE}
data <- read_excel("data/income.xls")
```

<p>&nbsp;</p>


### **Data manipulation**

```{r message=FALSE, warning=FALSE}
#Get dataset between 4-10 rows
data %<>%
  slice(4:10)


#Transpose the dataframe
data <- as.data.frame(t(as.matrix(data)))


#Rename variables
names(data) <- c("year", "total", "illiterate",  "literate_with_no_degree",
                 "less_than_high_school", "high_school_or_equivalent",
                 "higher_education")


#Remove first row
data %<>%
  slice(2:nrow(data))


#Change the type of values from character to numeric
data <- data.frame(lapply(data, function(x) as.numeric(as.character(x))))


#Change factor levels and date format
data$year <- as.Date(as.character(data$year), format = "%Y")
```

<p>&nbsp;</p>


### **Structure of data**

```{r}
str(data)
```

<p>&nbsp;</p>


### **Change of income of different education levels over years line plot**

```{r message=FALSE, warning=FALSE}
data_melted <- melt(data, "year")
area_plot <- ggplot(data_melted, aes(x = year, y = value)) + 
  geom_line(aes(color = variable), lwd = 2) + 
  scale_x_continuous(breaks = data$year) +
  theme_wsj() +
  scale_colour_wsj() +
  theme(legend.text = element_text(size = 30),
        legend.title = element_blank(),
        axis.text = element_text(size = 25),
        axis.text.x = element_text(angle = 70, vjust = 0.5, face = "bold"),
        axis.title = element_text(size = 25, face = "bold"),
        axis.title.x = element_blank())
  anim <- area_plot +
    transition_reveal(year) +
    labs(x = "variable", y = "value")
    animate(anim + ease_aes(x = "linear"), width = 1500, height = 700, res = 50,
            nframes = 80)
```

<p>&nbsp;</p>
<p>&nbsp;</p>


# **Education Expenditure**

<p>&nbsp;</p>


### **Import the data**

```{r message=FALSE, warning=FALSE}
data <- read_excel("data/education_expenditure_per_student.xls")
```

<p>&nbsp;</p>


### **Data manipulation**

```{r message=FALSE, warning=FALSE}
#Get data between rows 5 and 63 and delete last column which shows expenditures
#in dollar terms
data <- slice(data, 5:63)
data <- select(data, -3,-5)


#Rename variables
names(data) <- c("level_of_education", "year",
                 "education_expenditure_per_student_TL")



#Remove unwanted characters
data$year <- gsub(".(r).", "", data$year)
data$level_of_education <- gsub(".(1).", "", data$level_of_education)


#Remove rows all values NA
data <- filter(data, !is.na(education_expenditure_per_student_TL))


#Change type to numeric
data <- data.frame(lapply(data, function(x) as.numeric(as.character(x))))


#Reorganize our first column
data$level_of_education <- NA
levels <- c("total", "pre-primary_education", "primary_school", "lower_secondary_school",
            "upper_secondary_education", "tertiary_education")

k <- 1
for(i in seq(1, nrow(data), 9)){
  data$level_of_education[i] <- levels[k]
  k <- k + 1
}

data %<>% fill(level_of_education)



#Convert level of education to columns
data <- dcast(data,  year ~ level_of_education)


#Reorder the columns
data <- cbind(year = data$year, data[, levels])
```

<p>&nbsp;</p>


### **Structure of data**

```{r}
str(data)
```

<p>&nbsp;</p>


### **Change of expenditures per student over years line plot**

```{r fig.height=6, fig.width=13, message=FALSE, warning=FALSE}
data_melted <- melt(data, "year")
change_plot <- ggplot(data_melted, aes(x = year, y = value)) + 
  geom_line(aes(color = variable), lwd=2) + 
  scale_x_continuous(breaks = data$year) +
  theme_bw()+  
  ylab("Turkish Liras") + 
  ggtitle("Expenditure Per Student in Turkey") +
  theme(legend.text = element_text(size = 10),
        legend.title = element_blank(),
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 15, face = "bold"),
        axis.text.x = element_text(angle = 50, vjust = 0.5),
        axis.title.y = element_text(size = 15, face = "bold"))

change_plot
```

<p>&nbsp;</p>
<p>&nbsp;</p>


# **Education Levels by Provinces**

<p>&nbsp;</p>


### **Import the data**

```{r message=FALSE, warning=FALSE}
data <- read_excel("data/cities.xls")
spatial <- st_read("data/tur_polbnda_adm1.shp")
```

<p>&nbsp;</p>


### **Data manipulation**

```{r message=FALSE, warning=FALSE}
#Set column names as row 5 with distinct numbers to make ready to delete
#male and female columns
names(data) <- paste(data[5, ], seq(1, ncol(data)))


#Drop male and female columns
data <- select(data, -contains("Male"), -contains("Female"))


#Make column names to values in row 3 in order to delete NA columns easily
names(data) <- data[3, ]


#Rename 3rd and 5th variable
names(data)[c(3,5)] <- c("province_code", "province_name")


#Drop NA columns
data <- select(data, -starts_with("NA"))


#Remove rows that contain total Turkey information at the first row of each year
#It is possible to do that by removing just NA rows thanks to Province code
#columns of corresponding rows filled with NA
data <- na.omit(data)


#We will calculate the weighted average of populations of each education level
#for all cities. At the plotting process we will use this value to color the map
edu_levels <- data.frame(lapply(select(data, -c(1:4, 14)), function(x) as.numeric(as.character(x))))
weight <- seq(0.1, 0.9, 0.1)
data$value <- apply(edu_levels, 1, weighted.mean, weight)


#Rename variables
new_names <- c("illiterate", "literate_without_diploma", "primary_school",
              "primary_education", "junior_and_vocational_high_school",
              "high_and_vocational_high_school", "universities", "master",
              "doctorate")
names(data)[5:13] <- new_names


#Create a column that contains all information of cities. We will use it
#on the graph by showing the information when hovering on the city
for (column in new_names)
  data[[column]] <- paste(column, data[[column]], sep = ": ")

data <- transform(data, info = paste(illiterate, literate_without_diploma,
                                     primary_school, primary_education,
                                     junior_and_vocational_high_school,
                                     high_and_vocational_high_school,
                                     universities, master, doctorate,
                                     sep = "\n"))


#Keep only year, province name, value and edu_levels columns
data <- select(data, c(1, 3, 15, 16))


#Change the data type of year column and rename it
data[, 1] <- data.frame(lapply(select(data, 1), function(x) as.numeric(as.character(x))))
names(data)[1] <- "year"


#Clear extra spaces from province names
data$province_name <- str_trim(data$province_name, side = c("both", "left", "right"))


#There is province code column in the spatial dataset, we need to obtain last two character
#which contains numeric code (e.g TUR005 --> 05)
spatial$province_code <- str_sub(spatial$adm1, start = 5)
spatial$province_code <- as.numeric(spatial$province_code)


#Set a column that contains row numbers to make our datasets eligible to join
data$province_code <- rep(1:81, 12)


#Joining datasets on province codes
df = left_join(data, spatial, by = "province_code")


#Update info column by adding city names
df <- transform(df, info = paste(adm1_en, info, sep = "\n"))


#Change our dataset type from dataframe to tibble to be able to plot
df <- as_tibble(df)
```

<p>&nbsp;</p>


### **Structure of data**

```{r}
str(df)
```

<p>&nbsp;</p>


### **Optional**
```{r}
#We will plot the graphics of the last 7 years for the purpose of keep the file size small. One who is 
#curious about prior years graphics may delete this chunk before running the code.
start <- nrow(df) * 5 / 12 + 1
end <- nrow(df)
df <- slice(df, start:end)
```

<p>&nbsp;</p>


### **Choropleth map of education by provinces**

```{r fig.height=7, fig.width=10, message=FALSE, warning=FALSE}
mp <- ggplot() +
  geom_sf(data = df, aes(fill = value,
                         geometry = geometry,
                         text = info,
                         frame = year)) +
  scale_fill_gradientn(colours = c("gray0", "green", "white"),
                       values = scales::rescale(c(-0.5, 0, 0.5)),
                       breaks=c(min(df$value), max(df$value)),
                       labels=c("Low", "High")) +
  theme_map() +
  labs(fill = element_blank()) +
  ggtitle("Education Level by Province in Turkey") +
  theme(plot.title = element_text(size = 14, face = "bold.italic"))
  
  
mp <- ggplotly(mp, tooltip = c("text", "frame")) 
mp <- style(mp, hoverlabel = list(bgcolor = "white"))


#Change hoveron settings manually because style(hoveron="fill") command does not
#work with multiple frames
for (frame in seq(1:length(mp$x$frames))){
  for (trace in seq(1:length(mp$x$frames[[frame]]$data))){
    mp$x$frames[[frame]]$data[[trace]]$hoveron <- "fill"
  }
}


mp
```




